---
title: "Calculating Drug Synergy Combinations with Replicates"
author: "Kieran Redpath"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
* You can run code from a .Rmd document using "Ctrl + Enter" ("Cmd + Return" on mac) on the line or a highlighted block of text. Click the arrow in the top right corner of a chunk to run the entire chunk.
* Ensure all packages are correctly installed (see "Required packages").
* Substitute the default values under the heading "Input" with your desired values.
* Click the "Knit" button at the top of your RStudio window.

### Required packages
* Only run this chunk of code the first time you use synergyfinder on your computer, but first make sure you delete a single # from the start of each line.
* By default this code won't be included in the output document.
```{r, include = F, echo = F}
## Run these lines if the relevant package is not installed (remove the "#" first though)

## dplyr and ggplot2
# install.packages(c("dplyr", "ggplot2"))

## gridExtra
# install.packages("gridExtra")

## openxlsx
# install.packages("openxlsx", dependencies = TRUE)

## synergyfinder
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("synergyfinder")
```

### Load packages and data
```{r, message = F}
# Load packages
library(dplyr)
library(ggplot2)
library(gridExtra)
library(openxlsx)
library(synergyfinder)
library(tidyverse)
library(cowplot)
```

## Input
* Edit this chunk to define the names of variables, including:
* file in read.table should be the name of the file containing your data, which should be moved into this folder
* synergy.type should be your desired synergy calculation
* data.type should be how your data is expressed
```{r}
# Load combination data for any number of correctly indexed drug combinations (see the template file for what this should look like).
# replace filename with a .xlsx file containing your data in the right format
synergy.frame <- read.xlsx("SynergyFinderData.xlsx")
synergy.frame <- na.omit(synergy.frame)

# Define which type of synergy score you want to use ("HSA", "Loewe", "Bliss", or "ZIP")
synergy.type = "Loewe"

# Define if your data is expressed as "viability" or "inhibition". Either way it should be as a percentage, not a decimal (i.e. out of 100, not 1)
data.type = "viability"
```

### Process data
```{r, warning = F, include = F}
# Set the random seed so results are reproducible
set.seed(42)

# Process the synergy data (check that data.type is correct for your data, accepts "inhibition" or "viability")
dose.response.mat <- ReshapeData(synergy.frame,
                                 data_type = data.type,
                                 impute = T,
                                 noise = F,
                                 iteration = 100
)
```

### Plot dose response curves
* Also saves these as .png's
```{r, warning = F}
# Plots curves within RStudio and the generated markdown document, but also saves a png in the current working directory
# PlotDoseResponse(dose.response.mat, block_ids = dose.response.mat[["drug_pairs"]][["block_id"]], save_file = TRUE,  summary_statistic = NULL, file_type = "png")
```

### Calculate and visualise synergy scores across the inferred dose-response matrix
* Also saves these as .png's
```{r, warning = F}
# Calculate synergy scores
synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = synergy.type,
                                  iteration = 100)

# Visualise synergy scores, providing an average value and a visual indication of where synergy is high and low
# Plots scores within RStudio and the generated markdown document, but also saves a png in the current working directory
# PlotSynergy(synergy.score, type = "2D", method = synergy.type, block_ids = dose.response.mat[["drug_pairs"]][["block_id"]], summary_statistic = NULL, save_file = TRUE, file_type = "png")
```

### Extract data
* This chunk of code extracts useful values from the data for visualization and plotting, for only the tested concentrations of each drug
* Data for each combo is saved in a different element of the list "synergy.response.data"
* By default the code won't be included in the output document because it's very long and complicated
```{r, include = F, echo = F}
# Create an empty list to enter data into
synergy.response.data <- list()
# Extract rows of values you've actually tested in the lab: removing the imputed values where n = 1
# First extract response information (calculated mean from replicates and other statistics)
tmp <- synergy.score[["response_statistics"]][which(synergy.score[["response_statistics"]]$n > 1),]
# Remove NAs - not sure if this line is necessary (or it might break stuff)
tmp <- na.omit(synergy.score[["response_statistics"]])
# Combine this with synergy scores (calculated mean and other statistics)
tmp <- merge(tmp, synergy.score[["synergy_scores_statistics"]],
             by = c("block_id", "conc1", "conc2"),
             sort = F)

# Add all of these values into a list that contains each combination/blocks data as a separate element
for(x in unique(tmp$block_id)){
  # Define the drug combination for this block
  block = synergy.score[["drug_pairs"]]$block_id[x]
  drug1 = synergy.score[["drug_pairs"]]$drug1[x]
  drug2 = synergy.score[["drug_pairs"]]$drug2[x]
  synergy.response.data[[x]] <- tmp[tmp$block_id == x,]
  
  # Move single drug response data into new columns
  # Create new columns for response and sem for drug 1
  synergy.response.data[[x]][, paste0(drug1, "_response")] <- NA
  synergy.response.data[[x]][, paste0(drug1, "_sem")] <- NA
  # Create new columns for response and sem for drug 2
  synergy.response.data[[x]][, paste0(drug2, "_response")] <- NA
  synergy.response.data[[x]][, paste0(drug2, "_sem")] <- NA
  
  # Move response values for drug 1 into a new column for readability
  # Define the value of y and remove 0's
  y <- unique(synergy.response.data[[x]]$conc1)
  y <- y[! y %in% "0"]
  # Move values from "response" to "drug1_response" based on conditions
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug1, "_response")][synergy.response.data[[x]]$conc1 == y[z]] <- synergy.response.data[[x]]$response_mean[synergy.response.data[[x]]$conc1 == y[z] & synergy.response.data[[x]]$conc2 == 0]
  }
  # Do the same for "response_sem"
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug1, "_sem")][synergy.response.data[[x]]$conc1 == y[z]] <- synergy.response.data[[x]]$response_sem[synergy.response.data[[x]]$conc1 == y[z] & synergy.response.data[[x]]$conc2 == 0]
  }
  
  # Move response values for drug 2 into a new column for readability
  # Define the value of y and remove 0's
  y <- unique(synergy.response.data[[x]]$conc2)
  y <- y[! y %in% "0"]
  # Move values from "response" to "drug2_response" based on conditions
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug2, "_response")][synergy.response.data[[x]]$conc2 == y[z]] <- synergy.response.data[[x]]$response_mean[synergy.response.data[[x]]$conc2 == y[z] & synergy.response.data[[x]]$conc1 == 0]
  }
  # Do the same for "response_sem"
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug2, "_sem")][synergy.response.data[[x]]$conc2 == y[z]] <- synergy.response.data[[x]]$response_sem[synergy.response.data[[x]]$conc2 == y[z] & synergy.response.data[[x]]$conc1 == 0]
  }
  
  # Remove extra rows
  synergy.response.data[[x]] <- synergy.response.data[[x]][!synergy.response.data[[x]]$conc1 == 0, ]
  synergy.response.data[[x]] <- synergy.response.data[[x]][!synergy.response.data[[x]]$conc2 == 0, ]
  
  # Define which columns you want to keep - currently this is designed for simplicity and only includes standard error of the mean (but this should probably actually be standard deviation)
  keep.columns <- c("conc1", "conc2",
                    paste0(drug1, "_response"), paste0(drug1, "_sem"),
                    paste0(drug2, "_response"), paste0(drug2, "_sem"),
                    "response_mean", "response_sem",
                    paste0(synergy.type, "_synergy_mean"), paste0(synergy.type, "_synergy_sem"))
  synergy.response.data[[x]] <- synergy.response.data[[x]][, keep.columns]
  
  # Rename rows for readability and plotting
  combo.names <- paste0(drug1, "_",
                        synergy.response.data[[x]]$conc1, dose.response.mat[["drug_pairs"]][["conc_unit1"]][x],
                        "_X_", drug2, "_",
                        synergy.response.data[[x]]$conc2, dose.response.mat[["drug_pairs"]][["conc_unit2"]][x])
  rownames(synergy.response.data[[x]]) <- combo.names
  # Rename some columns for readability and plotting
  synergy.response.data[[x]] <- synergy.response.data[[x]] %>% rename(
    !!paste0(drug1, "_conc") := "conc1",
    !!paste0(drug2, "_conc") := "conc2",
    "Combo_response" = "response_mean",
    "Combo_sem" = "response_sem")
  # Rename each list element
  names(synergy.response.data)[x] <- paste0(synergy.type, "_block", "_", x, "_", drug1, "_X_", drug2)
}
# Print or do further processing with new_data

# Clean up environment
rm("tmp","block", "drug1", "drug2", "combo.names", "keep.columns", "x", "y", "z")
```

### Save Data
* This chunk of code saves a single excel workbook, with a tab for each drug combination
```{r}
wb <- createWorkbook(title = "Synergy_Response_Data")
for(y in 1:length(synergy.response.data)){
  # Individual files
  # write.xlsx(synergy.response.data[[y]], file = paste0(names(synergy.response.data[y]), ".xlsx"), rowNames = T)
  # Workbook  
  prefix <- paste0(synergy.type, "_block_")
  sheet.name <- gsub(paste0("^", prefix, ".*?_"), "", names(synergy.response.data)[y])
  sheet.name <- gsub("_X_", "_", sheet.name)
  
  addWorksheet(wb, sheetName = sheet.name)
  writeData(wb, sheet = sheet.name, synergy.response.data[[y]])
}
saveWorkbook(wb, file = paste0(synergy.type, "_", "Synergy_Response_Data.xlsx"), overwrite = T)

# Clean up environment
rm("y", "wb", "prefix")
```

# Transform the synergy.response.data from SynergyFinder.Rmd back to viability (it's been treated converted to inhibition, we want it back to how we put it in) - note that these terms aren't exactly correct for organoid area, but same idea
* To do this, to every column in the list that ends in _response, apply 100-x
```{r}
# Apply the transformation to all data frames in the list
trans.synergy.response.data <- lapply(synergy.response.data, function(df) {
  # Identify columns ending in '_response'
  response_cols <- grep("_response$", colnames(df), value = TRUE)
  
  # Apply transformation 100 - x to those columns
  df[response_cols] <- lapply(df[response_cols], function(x) 100 - x)
  
  # Return the modified data frame
  return(df)
})
# 'transformed_synergy_response_data' now contains the modified data frames

saveRDS(trans.synergy.response.data, file = "viability_response_data.RData")
```

## Plotting

### Load data
```{r}
# Load raw data for plotting organoid area
# Define drugs
drugs <- c("AZD8055", "Defactinib", "Gefitinib")
# Create list object of drug data
drug.dat <- list()
for(i in drugs){
  # Load AZD8055 from master sheet copy in the project directory, which includes synergy info (added retrospectively)
  drug.dat[[i]] <- read.xlsx("Master_gSORG_Drugging_Template_AllReps.xlsx", paste0(i, "_R"))
}
```

### Setup data
```{r}
# Cut down to just the drug data, excluding the dasatinib and DKO organoid data. Also convert response to percentage
for(i in names(drug.dat)) {
  drug.dat[[i]] <- drug.dat[[i]] %>%
    filter(Drug == i | Drug == paste0("Dasatinib_", i)) %>%
    filter(Genotype == "WT" | Genotype == "CDH1") %>%
    mutate(DMSO_Normalised_Response_Mean = DMSO_Normalised_Response_Mean * 100)
}
```

### Define some valuable plot parameters
```{r}
# Quickly define variables for plotting to match the size and font of an A4 thesis page
A4.margin <- 2.54
A4.width <- 21 - 2 * A4.margin
A4.height <- 29.7 - 2 * A4.margin

# theme_set(theme_minimal(base_size = 12))
theme_set(theme_minimal(base_family = "serif"))
axis.title.size <- 12
axis.ticks.size <- 9
plot.title.size <- 12
```

### Plot synergy
```{r}
synergy.plot.list <- list()

# Plot each genotype
for(i in names(drug.dat)) {
  
  # Filter to just the first replicate per condition (which contains the synergy info)
  synergy.df <- drug.dat[[i]] %>%
    filter(Biological_Replicate == 1) %>%
    select(Genotype, Drug, Concentration_uM, Synergy, Synergy_SEM)
  
  # Get all genotypes present in this dataset
  genotypes <- unique(synergy.df$Genotype)
  
  # Create a separate plot for each genotype
  for (g in genotypes) {
    df_subset <- synergy.df %>% filter(Genotype == g)
    
    plot.name <- paste0(i, "_", g)  # Create a unique name for this plot
    
    synergy.plot.list[[plot.name]] <- ggplot(df_subset, aes(x = Concentration_uM, y = Synergy, color = Drug)) +
      geom_hline(yintercept = 0, col = "red", linetype = "longdash") +  # red dashed hline
      geom_ribbon(aes(ymin = Synergy - Synergy_SEM,
                      ymax = Synergy + Synergy_SEM),
                  alpha = 0.2, colour = NA) +  # transparent SEM ribbon
      geom_point(size = 1.3, colour = "grey50") +
      geom_line(aes(group = Drug), color = "grey50") +  # grey lines for all drugs
      # geom_errorbar(aes(ymin = Synergy - Synergy_SEM,
      #                   ymax = Synergy + Synergy_SEM), colour = "grey50",
      #               width = 0.05) +
      ylim(-15, 20) +
      ggtitle(ifelse(g == "WT", "Control", g)) +  # <- Add genotype title here
      labs(
        x = paste0(i, " concentration (μM)"),
        y = "Loewe synergy score \u00B1 SEM"
      ) +
      theme(
        plot.title = element_text(face = "bold", size = plot.title.size, hjust = 0.5),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = axis.ticks.size, colour = "black"),
        axis.title.y = element_text(size = axis.title.size, colour = "black", margin = margin(r = 5)),
        panel.grid.major = element_line(color = "grey85"),
        panel.grid.minor = element_line(color = "grey85"),
        plot.margin = margin(b = 5)
      )
  }
}
```

```{r}
synergy.plot.list <- list()

# Plot each genotype
for(i in names(drug.dat)) {
  
  # Filter to just the first replicate per condition (which contains the synergy info)
  synergy.df <- drug.dat[[i]] %>%
    filter(Biological_Replicate == 1) %>%
    select(Genotype, Drug, Concentration_uM, Synergy, Synergy_SEM)
  
  # Get all genotypes present in this dataset
  genotypes <- unique(synergy.df$Genotype)
  
  # Create a separate plot for each genotype
  for (g in genotypes) {
    df_subset <- synergy.df %>% filter(Genotype == g)
    
    plot.name <- paste0(i, "_", g)  # Create a unique name for this plot
    
    synergy.plot.list[[plot.name]] <- ggplot(df_subset, aes(x = Concentration_uM, y = Synergy)) +
      geom_hline(yintercept = 0, col = "red", linetype = "longdash") +
      geom_ribbon(aes(
        ymin = Synergy - Synergy_SEM,
        ymax = Synergy + Synergy_SEM,
        group = Drug
      ), fill = "grey50", alpha = 0.2) +
      geom_point(aes(group = Drug), size = 1.3, colour = "grey50") +
      geom_line(aes(group = Drug), colour = "grey50") +
      ylim(-15, 20) +
      ggtitle(
        if (g == "WT") {
          "Control"
        } else if (g == "CDH1") {
          bquote(bolditalic("Cdh1")^bolditalic("-/-"))
        } else {
          g
        }
      ) +
      
      labs(
        x = paste0(i, " concentration (μM)"),
        y = "Loewe synergy score \u00B1 SEM"
      ) +
      theme(
        plot.title = element_text(face = "bold", size = plot.title.size, hjust = 0.5),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size = axis.ticks.size, colour = "black"),
        axis.title.y = element_text(size = axis.title.size, colour = "black", margin = margin(r = 5)),
        panel.grid.major = element_line(color = "grey85"),
        panel.grid.minor = element_line(color = "grey85"),
        plot.margin = margin(b = 5)
      )
  }
}

```


### Plot organoid area
```{r}
# Create summary list
drug.dat.summary <- list()
# Create plot list
area.plot.list <- list()

# Plot each genotype
for(i in names(drug.dat)){
  # Summarize data to calculate mean and SEM for each genotype, drug, and Concentration_uM combination
  drug.dat.summary[[i]] <- drug.dat[[i]] %>%
    group_by(Genotype, Drug, Concentration_uM) %>%
    summarize(
      Mean_Response_All_Reps = mean(DMSO_Normalised_Response_Mean),  # Calculate the mean response
      SEM_All_Reps = sd(DMSO_Normalised_Response_Mean) / sqrt(n())   # Calculate the standard error of the mean (SEM)
    ) %>%
    ungroup()
  
  # Get all genotypes present in this dataset
  genotypes <- unique(drug.dat.summary[[i]]$Genotype)
  
  # Create a separate plot for each genotype
  for (g in genotypes) {
    df_subset <- drug.dat.summary[[i]] %>%
      filter(Genotype == g) %>%
      mutate(Drug = ifelse(Drug == paste0("Dasatinib_", i), paste0("Dasatinib (0.25µM) + ", i), Drug))
    
    # Ensure "Dasatinib + <i>" comes first in the legend
    df_subset$Drug <- factor(df_subset$Drug, levels = c(paste0("Dasatinib (0.25µM) + ", i), i))
    
    plot.name <- paste0(i, "_", g)  # Create a unique name for this plot
    
    # Define custom colors: current drug `i` = cyan, combo = darkblue
    color_vals <- setNames(
      object = c("darkblue", "cyan"),
      nm = c(paste0("Dasatinib (0.25µM) + ", i), i)
    )
    
    # Split the data by drug to control draw order
    df_list <- split(df_subset, df_subset$Drug)
    
    # Create base plot
    p <- ggplot() +
      coord_cartesian(ylim = c(0, 100)) +
      labs(
        x = paste0(i, " concentration (μM)"),
        y = "Relative organoid area \u00B1 SEM (%)",
        title = NULL
      ) +
      theme(
        text = element_text(colour = "black"),
        axis.title.x = element_blank(),
        legend.text = element_text(size = axis.ticks.size, colour = "black"),
        legend.title = element_text(size = axis.title.size, colour = "black"),
        panel.grid.major = element_line(color = "grey85"),
        panel.grid.minor = element_line(color = "grey85"),
        axis.text.y = element_text(size = axis.ticks.size, colour = "black"),
        axis.title.y = element_text(size = axis.title.size, colour = "black", margin = margin(r = 5))
      )
    
    # Add geoms grouped by Drug, in desired draw order
    for (drug_name in levels(df_subset$Drug)) {
      p <- p +
        geom_line(data = df_list[[drug_name]], aes(x = Concentration_uM, y = Mean_Response_All_Reps, group = Drug, color = Drug)) +
        geom_point(data = df_list[[drug_name]], aes(x = Concentration_uM, y = Mean_Response_All_Reps, color = Drug), size = 1.3) +
        geom_ribbon(data = df_list[[drug_name]],
                    aes(x = Concentration_uM,
                        ymin = Mean_Response_All_Reps - SEM_All_Reps,
                        ymax = Mean_Response_All_Reps + SEM_All_Reps,
                        fill = Drug), alpha = 0.2)
    }
    
    # Add manual color scale and fill color and legend title
    p <- p + scale_color_manual(values = color_vals, name = "Treatment") + scale_fill_manual(values = color_vals, name = "Treatment")
    
    # Save to list
    area.plot.list[[plot.name]] <- p
  }
}

# Tidy up the environment
rm(list = c("df_subset", "g", "i", "plot.name", "df_list", "drug_name", "color_vals"))

```

### Tile plots with cowplot::plot_grid
* Need to add legends to CDH1 plots, and remove y axis labels and ticks from them
* Need to somehow add a shared x axis label that covers both x axes
```{r}
# Create object to store combined plots
combined.plot <- list()

for(i in drugs){
  # Define names of plots to call
  panel.A <- synergy.plot.list[[paste0(i, "_WT")]] +
    theme(legend.position = "none")
  panel.B <- synergy.plot.list[[paste0(i, "_CDH1")]] + 
    theme(legend.position = "none", axis.title.y = element_blank())
  panel.C <- area.plot.list[[paste0(i, "_WT")]] + 
    theme(legend.position = "none")
  
  # Extract legend from panel.D *before* removing it
  legend.plot <- area.plot.list[[paste0(i, "_CDH1")]]
  legend <- cowplot::get_legend(legend.plot + theme(legend.position = "bottom"))
  
  # Now remove legend from panel.D for plotting
  panel.D <- legend.plot + 
    theme(legend.position = "none", axis.title.y = element_blank())
  
  # Create the shared title with bold and size 12 font
  shared.title <- ggdraw() + 
    draw_label(
      paste0("Dasatinib and ", i, " Synergy and Response in Gastric Organoids"),
      size = 12, fontface = "bold", hjust = 0.5, fontfamily = "Times New Roman"
    )
  
  # Combine the panels (A, B, C, D)
  combined.panels <- plot_grid(
    panel.A, panel.B, panel.C, panel.D,
    rel_heights = c(1.0, 1.0, 1.0, 1.0),
    nrow = 2, ncol = 2, axis = "tblr", align = "v", byrow = TRUE,
    labels = c("A", "B", "C", "D"), label_size = plot.title.size
  )
  
  # Add the shared title and x-axis label
  combined.with.title <- plot_grid(
    shared.title,
    combined.panels,
    ggdraw() + draw_label(
      paste0(i, " concentration (µM)"),
      size = axis.title.size,
      hjust = 0.5,
      fontfamily = "Times New Roman"
    ),
    ncol = 1,
    rel_heights = c(0.1, 1, 0.03)  # Adjust heights accordingly for title and labels
  )
  
  # Add legend at the bottom
  combined.plot[[i]] <- plot_grid(
    combined.with.title,
    legend,
    ncol = 1,
    rel_heights = c(1, 0.08)  # Adjust height of legend as needed
  )
  
  # Set background to white and save
  combined.plot[[i]] <- combined.plot[[i]] +
    theme(plot.background = element_rect(fill = "white", color = NA))
  
  # Save the plot
  ggsave(paste0("Thesis_Plots/", i, ".png"),
         plot = combined.plot[[i]],
         width = A4.width, height = 20, units = "cm", dpi = 600)
  
  print(combined.plot[[i]])
}

# Tidy up the environment
rm(list = c("panel.A", "panel.B", "panel.C", "panel.D",
            "legend", "legend.plot", "shared.title",
            "combined.with.title"))
```

# Drug interaction models
## Run interaction models where intercept is CDH1 and combination therapy - better for interpreting effects of individual elements
## This version is saved.
```{r}
# Create empty lists to store model summaries
summary_3way <- list()
summary_2way <- list()

# Loop through each element of the drug.dat list
for(i in names(drug.dat)) {
  # New filtering and renaming
  drug.subset <- drug.dat[[i]] %>%
    mutate(
      Genotype = case_when(
        Genotype == "WT" ~ "B_Control",
        Genotype == "CDH1" ~ "A_CDH1",
        Genotype == "DKO" ~ "C_DKO"),
      Drug = case_when(
        Drug == i ~ paste0("B_", i),
        Drug == paste0("Dasatinib_", i) ~ paste0("A_Dasatinib_", i)
      )
    ) %>%
    filter(
      Drug %in% c(paste0("B_", i), paste0("A_Dasatinib_", i)),
      Genotype %in% c("B_Control", "A_CDH1")
    )
  
  # Plot data for each drug
  drug.subset %>% ggplot(., aes(x = Concentration_uM, 
                                y = DMSO_Normalised_Response_Mean, col = Drug)) +
    geom_jitter(width = 0.1, size = 2) + 
    facet_wrap(~Genotype) + 
    ggtitle(paste("Plot for", i))  # Add a title with the drug name
  
  # Fit three-way interaction model for each drug
  fit_3way_intn = lm(DMSO_Normalised_Response_Mean ~ Drug * Genotype * Concentration_uM, 
                     data = drug.subset)
  
  # Test residuals for normality formally
  print(paste("Shapiro Test for residuals (3-way interaction) for", i))
  print(shapiro.test(fit_3way_intn$residuals))
  
  # Plot residuals to check for normality visually
  hist(fit_3way_intn$residuals, main = paste("Residuals Histogram (3-way interaction)", i), 
       xlab = "Residuals", col = "gray")
  
  # Diagnostic plots for the 3-way interaction model
  par(mfrow = c(1, 2))
  plot(fit_3way_intn, 1)
  plot(fit_3way_intn, 2)
  
  # Summary of model for the 3-way interaction
  print(paste("Summary for 3-way interaction model for", i))
  print(summary(fit_3way_intn))
  
  # Save the coefficients summary for the 3-way interaction model in the list
  summary_3way[[i]] <- summary(fit_3way_intn)$coefficients
  
  # Fit two-way interaction model for each drug
  fit_partial_intn = lm(DMSO_Normalised_Response_Mean ~ Drug * Genotype + Concentration_uM, 
                        data = drug.subset)
  
  # Test residuals for normality formally
  print(paste("Shapiro Test for residuals (2-way interaction) for", i))
  print(shapiro.test(fit_partial_intn$residuals))
  
  # Plot residuals to check for normality visually
  hist(fit_partial_intn$residuals, main = paste("Residuals Histogram (2-way interaction)", i), 
       xlab = "Residuals", col = "gray")
  
  # Diagnostic plots for the 2-way interaction model
  par(mfrow = c(1, 2))
  plot(fit_partial_intn, 1)
  plot(fit_partial_intn, 2)
  
  # Summary of model for the 2-way interaction
  print(paste("Summary for 2-way interaction model for", i))
  print(summary(fit_partial_intn))
  
  # Save the coefficients summary for the 2-way interaction model in the list
  summary_2way[[i]] <- summary(fit_partial_intn)$coefficients
}

# Reset graphical parameters
par(mfrow = c(1, 1))

# Now, `summary_3way` contains the coefficients summaries for the 3-way models for all drugs
# And `summary_2way` contains the coefficients summaries for the 2-way models for all drugs

View(summary_2way[["AZD8055"]])
View(summary_2way[["Defactinib"]])
View(summary_2way[["Gefitinib"]])


# Create a new workbook
wb <- createWorkbook()
# Save as individual csv's and an excel workbook
for(i in names(summary_2way)){
  write.csv(summary_2way[[i]], file = paste0("Interactions/", i, "_2Way.csv"))
  # Add a new worksheet for each drug
  addWorksheet(wb, sheetName = i)
  # Write the corresponding summary table to the worksheet
  writeData(wb, sheet = i, summary_2way[[i]], rowNames = T)
}
# Save the workbook to a file
saveWorkbook(wb, "Interactions/summary_2way_interactions.xlsx", overwrite = T)

# Tidy up the environment
rm(list = c("wb"))
```

## Run interaction models where intercept is CDH1 and combination therapy - better for interpreting effects of individual elements
## This version is saved.
```{r}
# Create empty lists to store model summaries
summary_3way <- list()
summary_2way <- list()

# Loop through each element of the drug.dat list
for(i in names(drug.dat)) {
  # New filtering and renaming
  drug.subset <- drug.dat[[i]] %>%
    mutate(
      Genotype = case_when(
        Genotype == "WT" ~ "A_Control",
        Genotype == "CDH1" ~ "B_CDH1",
        Genotype == "DKO" ~ "C_DKO"),
      Drug = case_when(
        Drug == i ~ paste0("A_", i),
        Drug == paste0("Dasatinib_", i) ~ paste0("B_Dasatinib_", i)
      )
    ) %>%
    filter(
      Drug %in% c(paste0("A_", i), paste0("B_Dasatinib_", i)),
      Genotype %in% c("A_Control", "B_CDH1")
    )
  
  # Plot data for each drug
  drug.subset %>% ggplot(., aes(x = Concentration_uM, 
                                y = DMSO_Normalised_Response_Mean, col = Drug)) +
    geom_jitter(width = 0.1, size = 2) + 
    facet_wrap(~Genotype) + 
    ggtitle(paste("Plot for", i))  # Add a title with the drug name
  
  # Fit three-way interaction model for each drug
  fit_3way_intn = lm(DMSO_Normalised_Response_Mean ~ Drug * Genotype * Concentration_uM, 
                     data = drug.subset)
  
  # Test residuals for normality formally
  print(paste("Shapiro Test for residuals (3-way interaction) for", i))
  print(shapiro.test(fit_3way_intn$residuals))
  
  # Plot residuals to check for normality visually
  hist(fit_3way_intn$residuals, main = paste("Residuals Histogram (3-way interaction)", i), 
       xlab = "Residuals", col = "gray")
  
  # Diagnostic plots for the 3-way interaction model
  par(mfrow = c(1, 2))
  plot(fit_3way_intn, 1)
  plot(fit_3way_intn, 2)
  
  # Summary of model for the 3-way interaction
  print(paste("Summary for 3-way interaction model for", i))
  print(summary(fit_3way_intn))
  
  # Save the coefficients summary for the 3-way interaction model in the list
  summary_3way[[i]] <- summary(fit_3way_intn)$coefficients
  
  # Fit two-way interaction model for each drug
  fit_partial_intn = lm(DMSO_Normalised_Response_Mean ~ Drug * Genotype + Concentration_uM, 
                        data = drug.subset)
  
  # Test residuals for normality formally
  print(paste("Shapiro Test for residuals (2-way interaction) for", i))
  print(shapiro.test(fit_partial_intn$residuals))
  
  # Plot residuals to check for normality visually
  hist(fit_partial_intn$residuals, main = paste("Residuals Histogram (2-way interaction)", i), 
       xlab = "Residuals", col = "gray")
  
  # Diagnostic plots for the 2-way interaction model
  par(mfrow = c(1, 2))
  plot(fit_partial_intn, 1)
  plot(fit_partial_intn, 2)
  
  # Summary of model for the 2-way interaction
  print(paste("Summary for 2-way interaction model for", i))
  print(summary(fit_partial_intn))
  
  # Save the coefficients summary for the 2-way interaction model in the list
  summary_2way[[i]] <- summary(fit_partial_intn)$coefficients
}

# Reset graphical parameters
par(mfrow = c(1, 1))

# Now, `summary_3way` contains the coefficients summaries for the 3-way models for all drugs
# And `summary_2way` contains the coefficients summaries for the 2-way models for all drugs

View(summary_2way[["AZD8055"]])
View(summary_2way[["Defactinib"]])
View(summary_2way[["Gefitinib"]])


# Create a new workbook
wb <- createWorkbook()
# Save as individual csv's and an excel workbook
for(i in names(summary_2way)){
  write.csv(summary_2way[[i]], file = paste0("Interactions/", i, "_2Way_B.csv"))
  # Add a new worksheet for each drug
  addWorksheet(wb, sheetName = i)
  # Write the corresponding summary table to the worksheet
  writeData(wb, sheet = i, summary_2way[[i]], rowNames = T)
}
# Save the workbook to a file
saveWorkbook(wb, "Interactions/summary_2way_interactions_B.xlsx", overwrite = T)

# Tidy up the environment
rm(list = c("wb"))
```

### Interpretation of three-way model
* DrugDefactinib: overall, there is a significant difference between the two drugs (across genotype and concentrarion level) in terms of their effect on organoid area (positive coefficient means oganoid area higher for Defactinib compared to Dasatinib_Defactinib).
* Concentration_uM: there is a significant negative association between concentration and organoid area (across genotype and drug type) - as concentration increases, organoid area decreases.
* DrugDefactinib:GenotypeWT: there is a significant interaction between drug type and genotype in terms of organoid area - i.e., the difference between Defactinib and Dasatinib_Defactinib is dependent on genotype: this manifests as the organoid area being higher for Defactinib samples (relariev to Dasatinib_Defactinib) in the CDH1 genotype.
