---
title: "Calculating Drug Synergy Combinations with Replicates"
author: "Kieran Redpath"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(dplyr)
library(ggplot2)
library(gridExtra)
library(openxlsx)
library(synergyfinder)
library(cowplot)
```

## Input

* Edit this chunk to define the names of variables, including:
* File in read.table should be the name of the file containing your data, which should be moved into this folder
* synergy.type should be your desired synergy calculation
* data.type should be how your data is expressed
```{r}
# Load combination data for any number of correctly indexed drug combinations (see the template file for what this should look like).
# replace filename with a .xlsx file containing your data in the right format
synergy.frame <- read.csv("16_Drugs_Viability_Averages_cell_lines.csv")
synergy.frame <- na.omit(synergy.frame)

# Define which type of synergy score you want to use ("HSA", "Loewe", "Bliss", or "ZIP")
synergy.type = "Loewe"

# Define if your data is expressed as "inhibition" or "viability". Either way it should be as a percentage, not a decimal (i.e. out of 100, not 1)
data.type = "viability"
```

### Optionally rename some drugs with aliases in this data frame, for the entire rest of the analysis
```{r}
synergy.frame <- synergy.frame %>%
  mutate(drug_col = case_when(
    drug_col == "AZD3759" ~ "Zorifertinib",
    TRUE ~ drug_col  # Keep other values unchanged
  ))
```

### Process data
```{r, warning = F, include = F}
# Set the random seed so results are reproducible
set.seed(42)

# Process the synergy data (check that data.type is correct for your data, accepts "inhibition" or "viability")
dose.response.mat <- ReshapeData(synergy.frame,
                                 data_type = data.type,
                                 impute = T,
                                 noise = F,
                                 iteration = 100
)
```

### Calculate and visualise synergy scores across the inferred dose-response matrix
* Also saves these as .png's if required
```{r, warning = F}
# Calculate synergy scores
synergy.score <- CalculateSynergy(data = dose.response.mat,
                                  method = synergy.type,
                                  iteration = 100)
```

### Extract data
* This chunk of code extracts useful values from the data for visualization and plotting, for only the tested concentrations of each drug
* Data for each combo is saved in a different element of the list "synergy.response.data"
* By default the code won't be included in the output document because it's very long and complicated
```{r, include = F, echo = F}
# Create an empty list to enter data into
synergy.response.data <- list()
# Extract rows of values you've actually tested in the lab: removing the imputed values where n = 1
# First extract response information (calculated mean from replicates and other statistics)
tmp <- synergy.score[["response_statistics"]][which(synergy.score[["response_statistics"]]$n > 1),]
# Remove NAs - not sure if this line is necessary (or it might break stuff)
tmp <- na.omit(synergy.score[["response_statistics"]])
# Combine this with synergy scores (calculated mean and other statistics)
tmp <- merge(tmp, synergy.score[["synergy_scores_statistics"]],
             by = c("block_id", "conc1", "conc2"),
             sort = F)

# Add all of these values into a list that contains each combination/blocks data as a separate element
for(x in unique(tmp$block_id)){
  # Define the drug combination for this block
  block = synergy.score[["drug_pairs"]]$block_id[x]
  drug1 = synergy.score[["drug_pairs"]]$drug1[x]
  drug2 = synergy.score[["drug_pairs"]]$drug2[x]
  synergy.response.data[[x]] <- tmp[tmp$block_id == x,]
  
  # Move single drug response data into new columns
  # Create new columns for response and sem for drug 1
  synergy.response.data[[x]][, paste0(drug1, "_response")] <- NA
  synergy.response.data[[x]][, paste0(drug1, "_sem")] <- NA
  # Create new columns for response and sem for drug 2
  synergy.response.data[[x]][, paste0(drug2, "_response")] <- NA
  synergy.response.data[[x]][, paste0(drug2, "_sem")] <- NA
  
  # Move response values for drug 1 into a new column for readability
  # Define the value of y and remove 0's
  y <- unique(synergy.response.data[[x]]$conc1)
  y <- y[! y %in% "0"]
  # Move values from "response" to "drug1_response" based on conditions
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug1, "_response")][synergy.response.data[[x]]$conc1 == y[z]] <- synergy.response.data[[x]]$response_mean[synergy.response.data[[x]]$conc1 == y[z] & synergy.response.data[[x]]$conc2 == 0]
  }
  # Do the same for "response_sem"
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug1, "_sem")][synergy.response.data[[x]]$conc1 == y[z]] <- synergy.response.data[[x]]$response_sem[synergy.response.data[[x]]$conc1 == y[z] & synergy.response.data[[x]]$conc2 == 0]
  }
  
  # Move response values for drug 2 into a new column for readability
  # Define the value of y and remove 0's
  y <- unique(synergy.response.data[[x]]$conc2)
  y <- y[! y %in% "0"]
  # Move values from "response" to "drug2_response" based on conditions
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug2, "_response")][synergy.response.data[[x]]$conc2 == y[z]] <- synergy.response.data[[x]]$response_mean[synergy.response.data[[x]]$conc2 == y[z] & synergy.response.data[[x]]$conc1 == 0]
  }
  # Do the same for "response_sem"
  for(z in 1:length(y)){
    synergy.response.data[[x]][, paste0(drug2, "_sem")][synergy.response.data[[x]]$conc2 == y[z]] <- synergy.response.data[[x]]$response_sem[synergy.response.data[[x]]$conc2 == y[z] & synergy.response.data[[x]]$conc1 == 0]
  }
  
  # Remove extra rows
  synergy.response.data[[x]] <- synergy.response.data[[x]][!synergy.response.data[[x]]$conc1 == 0, ]
  synergy.response.data[[x]] <- synergy.response.data[[x]][!synergy.response.data[[x]]$conc2 == 0, ]
  
  # Define which columns you want to keep - currently this is designed for simplicity and only includes standard error of the mean (but this should probably actually be standard deviation)
  keep.columns <- c("conc1", "conc2",
                    paste0(drug1, "_response"), paste0(drug1, "_sem"),
                    paste0(drug2, "_response"), paste0(drug2, "_sem"),
                    "response_mean", "response_sem",
                    paste0(synergy.type, "_synergy_mean"), paste0(synergy.type, "_synergy_sem"))
  synergy.response.data[[x]] <- synergy.response.data[[x]][, keep.columns]
  
  # Rename rows for readability and plotting
  combo.names <- paste0(drug1, "_",
                        synergy.response.data[[x]]$conc1, dose.response.mat[["drug_pairs"]][["conc_unit1"]][x],
                        "_X_", drug2, "_",
                        synergy.response.data[[x]]$conc2, dose.response.mat[["drug_pairs"]][["conc_unit2"]][x])
  rownames(synergy.response.data[[x]]) <- combo.names
  # Rename some columns for readability and plotting
  synergy.response.data[[x]] <- synergy.response.data[[x]] %>% rename(
    !!paste0(drug1, "_conc") := "conc1",
    !!paste0(drug2, "_conc") := "conc2",
    "Combo_response" = "response_mean",
    "Combo_sem" = "response_sem")
  # Rename each list element
  names(synergy.response.data)[x] <- paste0(synergy.type, "_block", "_", x, "_", drug1, "_X_", drug2)
}
# Print or do further processing with new_data

# Clean up environment
rm("tmp","block", "drug1", "drug2", "combo.names", "keep.columns", "x", "y", "z")
```

### Save Data
* This chunk of code saves excel files of the output for each drug combination, and a multi-tab workbook of all drugs together
```{r}
wb <- createWorkbook(title = "Synergy_Response_Data_cell_lines")
for(y in 1:length(synergy.response.data)){
  # Workbook
  prefix <- paste0(synergy.type, "_block_.*_Dasatinib_X_")
  sheet.name <- gsub(prefix, "", names(synergy.response.data)[y])
  addWorksheet(wb, sheetName = sheet.name)
  writeData(wb, sheet = sheet.name, synergy.response.data[[y]])
}
saveWorkbook(wb, file = paste0(synergy.type, "_", "Synergy_Response_Data_cell_lines.xlsx"), overwrite = T)

# Clean up environment
rm("y", "wb", "prefix")
```

## Plot Synergy, Inhibition, and Viability (on a log2 Scale)
### Transform the synergy.response.data from SynergyFinder.Rmd back to viability (it's been treated converted to inhibition, we want it back to how we put it in)
* To do this, to every column in the list that ends in _response, apply 100-x
```{r}
# Apply the transformation to all data frames in the list
trans.synergy.response.data <- lapply(synergy.response.data, function(df) {
  # Identify columns ending in '_response'
  response_cols <- grep("_response$", colnames(df), value = TRUE)
  
  # Apply transformation 100 - x to those columns
  df[response_cols] <- lapply(df[response_cols], function(x) 100 - x)
  
  # Return the modified data frame
  return(df)
})
# 'trans.synergy.response.data' now contains the modified data frames
```

### Plot synergy and viability
```{r}
# Create a vector for plotting minor axis ticks
minor.ticks <- list()
for(n in 1:7){
  minor.ticks[[n]] <- c(n+0.5, n+0.75, n+0.875)
}
minor.ticks <- unlist(minor.ticks)

# Quickly define variables for plotting to match the size and font of an A4 page with standard margins
A4.margin <- 2.54
A4.width <- 21 - 2 * A4.margin
A4.height <- 29.7 - 2 * A4.margin

theme_set(theme_minimal(base_family = "serif"))
axis.title.size <- 12
axis.ticks.size <- 9
plot.title.size <- 12

for(i in 1:length(trans.synergy.response.data)){
  # Specify current drugs
  # Hard-coded columns 1 and 2
  drug1 <- gsub("_conc", "", colnames(trans.synergy.response.data[[i]])[1])
  drug2 <- gsub("_conc", "", colnames(trans.synergy.response.data[[i]])[2])
  
  # For Viability Plot
  # Set up dataframe of relevant data for viability plot
  v.df <- as.data.frame(trans.synergy.response.data[[i]])[,1:8]
  v.df <- v.df %>% mutate_if(is.numeric, round, digits = 3) 
  v.df$dose <- rownames(v.df)
  v.df$data_order <- seq(1, nrow(v.df))
  v.df$exp_order <- 2^(1:nrow(v.df))
  
  # For Synergy Plot
  # Set up dataframe of relevant data for synergy plot
  s.df <- as.data.frame(trans.synergy.response.data[[i]])[,9:10]
  s.df <- s.df %>% mutate_if(is.numeric, round, digits = 3) 
  s.df$dose <- rownames(s.df)
  s.df$data_order <- seq(1, nrow(s.df))

  # For viability plot
  labels <- c()
  for(j in 1:nrow(v.df)){
    labels[j] <- paste0(v.df[j,2])
  }
  v.df$labels <- labels
  
  # For synergy plot
  labels  <- c()
  for(j in 1:nrow(v.df)){
    labels[j] <- paste0(v.df[j,1])
  }
  s.df$labels <- labels
  
  # Plot Synergy
  synergy.plot <- ggplot(data = s.df, mapping = aes(x = v.df$data_order, y = s.df[,paste0(synergy.type, "_synergy_mean")])) +
    geom_blank(data = s.df, mapping = aes(x = v.df$data_order, y = s.df[,paste0(synergy.type, "_synergy_mean")])) +
    geom_hline(yintercept = 0, col = "red", linetype = "longdash") +
    geom_point(mapping = aes(x = v.df$data_order, y = s.df[, paste0(synergy.type, "_synergy_mean")]), size = 1.3, colour = "grey50") +
    geom_line(mapping = aes(x = v.df$data_order, y = s.df[, paste0(synergy.type, "_synergy_mean")]), group = 1, colour = "grey50") +
    geom_ribbon(aes(ymin = s.df[, paste0(synergy.type, "_synergy_mean")] - s.df[, paste0(synergy.type, "_synergy_sem")],
                    ymax = s.df[, paste0(synergy.type, "_synergy_mean")] + s.df[, paste0(synergy.type, "_synergy_sem")]),
                fill = "grey50", alpha = 0.2) +
    coord_cartesian(ylim = c(-30, 20)) +
    scale_x_continuous(breaks = v.df$data_order, minor_breaks = minor.ticks, expand = expansion(add = 0.25)) +
    theme(title = element_text(size = plot.title.size, margin = margin(t = 30)),
          axis.text.x = element_blank(), axis.title.x = element_blank(), plot.title = element_text(hjust = 0.5, colour = "black", face = "bold", size = plot.title.size), axis.text.y = element_text(size = axis.ticks.size, colour = "black"), axis.title.y = element_text(size = axis.title.size, colour = "black", margin = margin(r = 5)), panel.grid.major = element_line(color = "grey85"), panel.grid.minor = element_line(color = "grey85"),
          plot.margin = margin(b = 5)) +
    ylab(paste0(synergy.type, " synergy score ± SEM")) +
    ggtitle(paste0(drug1, " and ", drug2, " Synergy and Viability in NCI-N87"))
  
  # Plot Viability
  viability.plot <- ggplot(data = v.df) +
    
    # Plot lines, points, and ribbons
    
    # For combination
    geom_line(mapping = aes(x = v.df$data_order, y = Combo_response, group = 1, colour = paste0(drug1, " + ", drug2))) +
    geom_point(mapping = aes(x = v.df$data_order, y = Combo_response, group = 1, colour = paste0(drug1, " + ", drug2)),
               size = 1.3) +
    geom_ribbon(mapping = aes(x = v.df$data_order, ymin = Combo_response - Combo_sem, ymax = Combo_response + Combo_sem, fill = paste0(drug1, " + ", drug2)), alpha = 0.2) +
    
    # For drug 1
    geom_line(mapping = aes(x = v.df$data_order, y = v.df[, paste0(drug1, "_response")], group = 1, colour = drug1)) +
    geom_point(mapping = aes(x = v.df$data_order, y = v.df[, paste0(drug1, "_response")], group = 1, colour = drug1), size = 1.3) +
    geom_ribbon(mapping = aes(x = v.df$data_order, ymin = v.df[, paste0(drug1, "_response")] - v.df[, paste0(drug1, "_sem")], ymax = v.df[, paste0(drug1, "_response")] + v.df[, paste0(drug1, "_sem")],  fill = drug1), alpha = 0.2) +
    
    # For drug 2
    geom_line(mapping = aes(x = v.df$data_order, y = v.df[, paste0(drug2, "_response")], group = 1, colour = drug2)) +
    geom_point(mapping = aes(x = v.df$data_order, y = v.df[, paste0(drug2, "_response")], group = 1, colour = drug2), size = 1.3) +
    geom_ribbon(mapping = aes(x = v.df$data_order,  ymin = v.df[, paste0(drug2, "_response")] - v.df[, paste0(drug2, "_sem")], ymax = v.df[, paste0(drug2, "_response")] + v.df[, paste0(drug2, "_sem")], fill = drug2), alpha = 0.2) +
    
    # Set colours
    scale_color_manual(name = "Treatment", values = c("darkblue", "black", "cyan"), limits = c(paste0(drug1, " + ", drug2), drug1, drug2)) +
    scale_fill_manual(name = "Treatment", values = c("darkblue", "black", "cyan"), limits = c(paste0(drug1, " + ", drug2), drug1, drug2)) +   
    
    # Provide titles and aesthetic parameters
    coord_cartesian(ylim=c(0, 120)) +
    scale_x_continuous(breaks = v.df$data_order, minor_breaks = minor.ticks, expand = expansion(add = 0.25)) +
    theme(text = element_text(colour = "black"), axis.text.x = element_blank(), axis.title.x = element_blank(),
          legend.text = element_text(size = axis.ticks.size, colour = "black"), legend.title = element_text(size = axis.title.size, colour = "black"), panel.grid.major = element_line(color = "grey85"), panel.grid.minor = element_line(color = "grey85"),
          # legend.position = c(0.95, 0.3),
          axis.text.y = element_text(size = axis.ticks.size, colour = "black"), axis.title.y = element_text(size = axis.title.size, colour = "black", margin = margin(r = 5))) +
    ylab("Relative cell viability ± SEM (%)")
  
  # Label plot
  # All drugs are in µM apart from metformin, which is in mM
  if(drug2 == "Metformin"){
    druglabs <- c(rep(paste0(drug1, " (µM)"), nrow(s.df)), rep(paste0(drug2, " (mM)"), nrow(v.df)))
  } else{
    druglabs <- c(rep(paste0(drug1, " (µM)"), nrow(s.df)), rep(paste0(drug2, " (µM)"), nrow(v.df)))
  }
  # Create a table of labels
  label.table <- data.frame(
    exp_order = v.df$data_order,
    labels = c(s.df$labels, v.df$labels),
    drug = druglabs 
  )
  
  # Reorder the rows so that drug2 is always on top
  drug2_label <- ifelse(drug2 == "Metformin", paste0(drug2, " (mM)"), paste0(drug2, " (µM)"))
  drug1_label <- paste0(drug1, " (µM)")
  
  # Reorder the data frame based on the drug column
  label.table <- label.table %>%
    arrange(factor(drug, levels = c(drug2_label, drug1_label)))
  
  # Convert the drug column to a factor with specified levels to enforce order in the plot
  label.table$drug <- factor(label.table$drug, levels = c(drug2_label, drug1_label))
  
  table.plot <- ggplot(label.table, aes(x = exp_order, y = drug)) +
    geom_text(aes(label = labels), size = axis.ticks.size/.pt, family = "serif") +
    theme_void() +
    theme(axis.text.y = element_text(size = axis.ticks.size, colour = "black", margin = margin(r = 10)), axis.title.x = element_blank(), text = element_text(family = "serif")) +
    coord_cartesian(clip = "off") +
    scale_x_continuous(expand = expansion(add = 0.25))
  
  # Extract and remove legends
  viability.plot.with.legend <- viability.plot + theme(legend.position = "bottom")
  shared.legend <- cowplot::get_legend(viability.plot.with.legend)
  viability.plot <- viability.plot + theme(legend.position = "none")
  synergy.plot <- synergy.plot + theme(legend.position = "none")
  # Put them together
  combined.plot <- plot_grid(
    plot_grid(synergy.plot, viability.plot, table.plot,
              rel_heights = c(1.0, 1.4, 0.3), ncol = 1,
              axis = "lr", align = "v", byrow = TRUE,
              labels = c("A", "B", ""), label_size = plot.title.size),
    shared.legend,
    ncol = 1,
    rel_heights = c(1, 0.1)  # Adjust height of legend row if needed
  )
  
  ggsave(paste0("Synergy_Viability_Plots_cell_lines/", names(trans.synergy.response.data[i]), ".png"), plot = combined.plot, width = A4.width, height = A4.width, units = "cm", dpi = 600)
  print(combined.plot)
}

# Clean up the environment
rm(list = c("A4.height", "A4.margin", "A4.width",
            "axis.ticks.size", "axis.title.size", "plot.title.size", "minor.ticks",
            "i", "j", "n",
            "drug1", "drug2", "druglabs", "labels",
            "s.df", "v.df", "label.table",
            "synergy.plot", "viability.plot", "table.plot", "combined.plot"))
```
